Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs: Online Supplement

### 2014-01-31 Revision/Supplementary/supplementary.tex
added revision

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Probability Aggregation in Time-Series}

@@

 \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
 \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
 \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}

\end{aug}

\begin{abstract}
Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.  This paper advances into unexplored areas of probability aggregation by considering a dynamic context in which experts can update their beliefs at random intervals. The updates occur very infrequently, resulting in a highly sparse dataset that cannot be modeled by standard time-series procedures. In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts over a period of two years on different subsets of 166 international political events. 
\end{abstract}

\begin{keyword}
\kwd{Probability Aggregation}
\kwd{Dynamic Linear Model}
\kwd{Hierarchical Modeling}
\kwd{Expert Forecast}
\kwd{Subjective Probability}
\kwd{Bias Estimation}
\kwd{Calibration}
\kwd{Time Series}
\end{keyword}
\end{frontmatter}

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. Their probability assessments are often evaluated and compared on \textit{calibration}, which measures how closely the frequency of event occurrence agrees with the assigned probabilities. For instance, the proportion of occurrences is 60\% for all those events to which a well-calibrated expert assigned a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts. In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.

Calibration by itself, however, is not sufficient for useful probability estimation. Consider a relatively stationary process, such as rain on different days in a given geographic region, where the observed frequency of occurrence in the last 10 years is 45\%. In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated. This assessment, however, can be made without any subject-matter expertise or special training. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. Therefore experts should aim to make probability assessments that are as far from the baseline as possible. The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy. From the decision-maker's perspective their advice is highly useful as it involves much information and very little risk. Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}). This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.

%One such observation is related to the aggregation of multiple probabilities. 
There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance. Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date. To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly. Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable. This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item borrowing strength across the hierarchy to analyze under- and overconfidence in different groups of experts,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations, such as expert disagreement and problem difficulty, and can be used to gain novel insight in the social sciences. 
\end{itemize}

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to our real-world geopolitical forecasting database. The paper concludes with a discussion on future research directions and model limitations.

\section{Geopolitical Forecasting Data}
\label{data}

\begin{table*}
\caption{Five-number summaries of our real-world data.}
\label{DataStats}
\begin{tabular}{m{2in}	lllllll }
\hline
Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
\# of Days a Question is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
\# of Experts per Question & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
\# Forecasts given by each Expert on a Question & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
\# Questions participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
\end{tabular}
\end{table*}

\begin{table*}
\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
\label{ExpertiseTable}
\begin{center}
\begin{tabular}{llllllll }
\hline
Expertise Level & 1 & 2 & 3 & 4 & 5\\
Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
\end{tabular}
\end{center}
\end{table*}

The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period during which the participating experts could update their forecasts as frequently as they liked without penalty. The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts. Given that a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.

The experts made updates on a very infrequent basis: the average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the resulting dataset is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the dataset and its collection see \citet{ungar2012good}.

\begin{figure}[h!]
\vspace{-2em}
\centering
\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-3.5em}

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
\label{Example1}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
\label{Example2}
}

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
\label{ExamplePlots}
\end{figure}

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively correlated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$. Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$. The error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.

The hidden state $X_{t,k}$ represents the aggregate logit-probability for the $k$th event given all the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. It is reasonable to assume that this probability forecast 
%that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
%an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
\begin{mydef}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
\end{mydef}
\begin{mydef}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}). 
\end{mydef}
\noindent
Even though it is unlikely that any single expert has access to all the available information, a large and diverse group of experts may share a considerable portion of the available information. The collective wisdom of the group therefore provides an attractive proxy for $\mathcal{F}_{t,k}$.

Given that the experts may believe in false information, hide their true beliefs, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$. In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from $X_{t,k}$. Therefore an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.  Due to the high sparsity of our data, estimating a bias term separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from $X_{t,k}$ is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

 The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Given that each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

\section{Model Estimation}
\label{identifiability}

The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model. This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimizational procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

\subsection{Sampling Step}
\label{sampling_step}
Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  In Section \ref{syntheticData} we provide a guideline for choosing the reference point. Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Given that this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling steps, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Given that for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint.

\subsection{Discussion}
If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}). To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where 
\begin{eqnarray}
\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
\end{eqnarray}
This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a linear regression model
\begin{eqnarray*}
\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k} + e_{t,k}
\end{eqnarray*}
with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$. If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero. Given that the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero. Therefore both the response and explanatory variables are approximately centered. This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule. If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter. This, however, reduces the interpretability of our model. Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form. One procedure to attain this is described in the beginning of Section \ref{realData}.

If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ is extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, is chosen as the base-case to ensure that the probabilities will sum to one at any given time point. Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}). Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case. This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.

%
%\section{Extension: Polychotomous Outcomes}
%If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must be extended to a vector of size $M-1$. One of the outcomes, e.g., the $M$th one, is chosen as the base-case.  This gives us a total of $M-1$ observed processes and one hidden process
%\begin{eqnarray*}
%\boldsymbol{Y}_{1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{1}X_{1,t,k} + \boldsymbol{v}_{1,t,k}\\
%\boldsymbol{Y}_{2,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{2}X_{2,t,k} + \boldsymbol{v}_{2,t,k}\\
%&\vdots&\\
%\boldsymbol{Y}_{M-1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{M-1}X_{M-1,t,k} + \boldsymbol{v}_{M-1,t, k}\\
%\boldsymbol{X}_{t,k} &=& \boldsymbol{\gamma}^T_k \boldsymbol{I}_{M-1} \boldsymbol{X}_{t-1,k} + \boldsymbol{w}_{t,k}
%\end{eqnarray*}
%where $\boldsymbol{X}_{t,k}$ is a $(M-1) \times 1$ matrix of calibrated logit-probabilities. Notice that each process has a separate bias vector. The error terms are independent and identically distributed multivariate normal random variables 
%\begin{eqnarray*}
%\boldsymbol{v}_{m, t, k} | \sigma^2_{j,k} &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k} \left(\boldsymbol{0}, \sigma^2_{j,k} \boldsymbol{I}_{I_k} \right)\\
%\boldsymbol{w}_{t,k} | \tau_k^2 &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{M-1} \left(\boldsymbol{0}, \tau_k^2 \boldsymbol{I}_{M-1} \right)
%\end{eqnarray*}
%On the left-hand side of observed process
%\begin{eqnarray*}
%\boldsymbol{Y}_{m,t,k} &=&  \left[ \begin{matrix} Y_{m,1,t,k} & Y_{m,2,t,k} & \dots & Y_{m,I_k,t,k} \end{matrix} \right]^T\\
%&=&  \left[ \begin{matrix} \log \left( \frac{p_{m,1,t,k}}{p_{M,1,t,k}} \right) &  \log \left( \frac{p_{m,2,t,k}}{p_{M,2,t,k}} \right) & \dots &  \log \left( \frac{p_{m,I_k,t}}{p_{M,I_k,t,k}} \right) \end{matrix} \right]^T
%\end{eqnarray*}
%is a $I_k \times 1$ matrix of the expert logit-probabilities for the $m$th outcome of the $k$th question at time $t$, and on the left-hand side of the hidden process
%\begin{eqnarray*}
%\boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
%\end{eqnarray*}
%is the matrix of sharp and calibrated logit-probabilities at time $t$. Choosing one of the outcomes as the base-case, ensures that the probabilities will sum to one at any given time point. Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models (see Section \ref{model}), the estimation can be done separately for each outcome. Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize to the multi-outcome case.

\section{Synthetic Data Results}
\label{syntheticData}

The goal in this section is to evaluate the ability of the \textit{Sample-And-Calibrate}-procedure to capture true parameter values. The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly news, and (ii) it is unclear how the dynamic part of the model can be used to produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.

To ensure that the hidden process for question $k$ is well-calibrated, we generate a path of standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 50 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$ when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). Our simulation study considers a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Forty synthetic datasets are generated for each combination of $\sigma^2$, $\beta$, and $K$ values. The SAC-procedure runs for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.

%$|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively. The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.

\begin{figure}[h!]
\vspace{-5.5em}
\centering
 \hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package

\vspace{0em}
\hspace{2.5em}Hidden Process
\vspace{-0.5em}

\subfigure{
\hspace{-1em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2}
\label{SyntheticSigma2}
}
\subfigure{
\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBeta}
\label{SyntheticBeta}
}
\subfigure{
\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticK}
\label{SyntheticK}
}

\vspace{0.5em}
\hspace{2.5em}Bias Vector
\vspace{-0.5em}

\subfigure{
\hspace{-1em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2bias}
\label{SyntheticSigma2b}
}
\subfigure{
\hspace{-1.5em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBetabias}
\label{SyntheticBetab}
}
\subfigure{
\hspace{-1.5em}   \includegraphics[width = 0.32\textwidth]{Figures/SyntheticKbias}
\label{SyntheticKb}
}

\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data. The top row presents the accuracy to capture the hidden process. The bottom row presents the accuracy to capture the bias vector. }
\label{Synthetic}
\end{figure}

>>>>

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs: Online Supplement}
\runtitle{Probability Aggregation in Time-Series: Online Supplement}

@@

% \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
% \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
% \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
% \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
%  
\end{aug}

\end{frontmatter}

This supplementary material accompanies the paper ``Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs''. It provides a technical description of the sampling step of the   SAC-algorithm.

\section{Technical Details of the Sampling Step}
\label{appendix}
